---
title: "HW10"
author: "Ben Rochford"
format: html
editor: visual
---

```{r}
library(tidyverse)
theme_set(theme_light(base_family = "Optima"))

library(modelsummary)
library(broom)
library(gt)

library(gssr)
gss18 <- gss_get_yr(2018) 

d <- gss18 |> 
  select(attend, polviews, cappun, degree) |> 
  haven::zap_missing() |> 
  mutate(across(!degree, haven::zap_labels)) |> 
  mutate(degree = haven::as_factor(degree)) |> 
  mutate(
    weekly = if_else(attend >= 7, 1L, 0L),
    polviews = polviews - 4,
    cappun = if_else(cappun == 1, 1L, 0L),
    ) |> 
  mutate(conservative = as.integer(polviews > 0)) |> 
  drop_na() 

d
```

## 10.3.1

```{r}
mat311 <- model.matrix(cappun ~ polviews + weekly + polviews:weekly, data = d)
head(mat311)
dim(mat311)
```

`cappun` is an outcome variable here, while `polviews` and `weekly` and their interaction are predictor variables. Model matrix dataframe creates a column for intercept and for each predictor, for each row of the data.

```{r}
mat312 <- model.matrix(cappun ~ polviews * weekly, data = d)
head(mat312)
dim(mat312)
```

Syntactically equivalent to the first.

```{r}
mat313 <- model.matrix(cappun ~ degree, data = d)
head(mat313)
dim(mat313)
```

Now we have columns for the 4 predictors derived from `degree` which has 4 possible values. As a result, the model matrix has columns for the intercept and the 4 predictors, with rows for each data point.

## 10.3.2

```{r}
mod1 <- glm(cappun ~ conservative + weekly, data = d, family = "binomial")
mod2 <- glm(cappun ~ conservative * weekly, data = d, family = "binomial")
mod3 <- glm(cappun ~ polviews + weekly, data = d, family = "binomial")
mod4 <- glm(cappun ~ polviews * weekly, data = d, family = "binomial")
```

```{r}
msummary(list(mod1, mod2, mod3, mod4), output = "gt") |> 
  opt_table_font(font = "Optima")
```

### Comparing with AIC and BIC

The models that used the more granular `polviews` variable as a predictor rather than `conservative` had markedly lower AIC and BIC values. The model with the lowest overall AIC score was model 4, which predicted `cappun` using `polviews`, `weekly`, and their interaction.

### Interpreting Intercepts

The log odds of someone supporting capital punishment when they ...

-   1: are not conservative and do not attend church weekly = 0.358

-   2: are not conservative and do not attend church weekly!! = 0.341

-   3: are politically moderate and do not attend church weekly = 0.666

-   4: are politically moderate and do not attend church weekly!! = 0.677

We see that these models are predicting in general that moderates have a higher log odds of supporting capital punishment than "non conservatives."

### Using mod4

to predict probability that a slightly conservative individual that attends weekly religious ceremonies favors capital punishment:

polviews == 1 and weekly == 1

```{r}
plogis(0.677 + 0.441*1 - 0.451*1 - 0.138)
```

## 10.3.3

### Transforming polviews

```{r}
d <- d |> 
  mutate(polviews2 = case_when(polviews < 0 ~ "liberal",
                               polviews == 0 ~ "moderate",
                               polviews > 0 ~ "conservative")) |> 
  mutate(polviews2 = haven::as_factor(polviews2))
```

### Fitting new reference categories

```{r}
d$polviews2 <- relevel(d$polviews2, ref = "liberal")
mod331 <- glm(cappun ~ polviews2 + weekly, data = d, family = "binomial")

d$polviews2 <- relevel(d$polviews2, ref = "moderate")
mod332 <- glm(cappun ~ polviews2 + weekly, data = d, family = "binomial")

d$polviews2 <- relevel(d$polviews2, ref = "conservative")
mod333 <- glm(cappun ~ polviews2 + weekly, data = d, family = "binomial")

msummary(list(mod331, mod332, mod333), gof_map = "none", output = "gt") |> 
  opt_table_font(font = "Optima") 
```

### Interpreting coefficients

-   1: point of reference is liberals. Moderates have a .894 higher log odds of supporting capital punishment, and conservatives have a 1.362 higher log odds of supporting capital punishment

-   2: point of reference is moderates. Conservatives have a .468 higher log odds of supporting capital punishment, while liberals have a .894 lower log odds of supporting capital punishment

-   3: point of reference is conservatives. Moderates have a .468 lower log odds of supporting capital punishment, while liberals have a 1.362 lower log odds of supporting capital punishment

In all models, those that attend religious services weekly have a .417 lower log odds of supporting capital punishment.

## 10.3.4

### Entries for saturated model column

```{r}
# reset order
d$polviews2 <- relevel(d$polviews2, ref = "moderate")
d$polviews2 <- relevel(d$polviews2, ref = "liberal")

d |> 
  group_by(weekly, polviews2) |> 
  summarize(p = mean(cappun))
```

### Entries for restricted model column

```{r}
mod34 <- glm(cappun ~ polviews2 + weekly, data = d, family = "binomial")

d34 <- with(d, data.frame(weekly = c(0, 0, 0, 1, 1, 1), polviews2 = c("liberal","moderate","conservative","liberal","moderate","conservative")))

d34$p <- predict(mod34, newdata = d34, type = "response")
d34
```

## 10.3.5

### Picking new GSS outcome and predictors

Predicting being okay with extramarital sex based on gossiping a little, being female, and having a college degree.

```{r}
d_new <- gss18 |> 
  select(mcsds6, sex, xmarsex, degree) |> 
  haven::zap_missing() |> 
  haven::zap_labels() |> 
  mutate(cheating_cool = if_else(xmarsex >= 3 & xmarsex < 8, 1L, 0L),
         female = if_else(sex == 2, 1L, 0L),
         gossip = if_else(mcsds6 == 1, 1L, 0L),
         college = if_else(degree >= 3, 1L, 0L)) |> 
  drop_na()

mod351 <- glm(cheating_cool ~ female + gossip, data = d_new, family = "binomial")
mod352 <- glm(cheating_cool ~ female + gossip + college, data = d_new, family = "binomial")

msummary(list(mod351, mod352), output = "gt") |> 
  opt_table_font(font = "Optima")
```

### Comparing with AIC and BIC

The AIC of the model that factors in college degree is very slightly lower, but the BIC of the model that does not factor in college degree is lower.

Since AIC provides a good approximation for out-of-sample deviance, we can infer that the model that factors in college degree would have better out-of-sample predictive accuracy.

### Relationship between outcomes and predictors

In terms of the coefficients:

-   1: being female results in a .301 lower log odds of being okay with extramarital sex, and being a gossip results in a .478 higher log odds of being okay with extramarital sex

-   2: being female results in a .299 lower log odds of being okay with extramarital sex, and being a gossip results in a .443 higher log odds of being okay with extramarital sex, and having a college degree results in a .365 higher log odds of being okay with extramarital sex
