---
title: "HW7"
author: "Ben Rochford"
format: html
editor: visual
---

```{r}
library(tidyverse, warn.conflicts = FALSE)
library(infer)
library(janitor) ## for convenient "table" functions
library(gssr)    ## for access to GSS data
library(ggthemes)
theme_set(theme_few(base_family = "Optima")) 
```

## 7.1.1

Load 2018 GSS data

```{r}
library(tidyverse)
library(gssr)
gss18 <- gss_get_yr(2018) 

d <- gss18 |> 
  select(attend, polviews) |> 
  haven::zap_missing() |> 
  haven::zap_labels() |> 
  mutate(weekly = if_else(attend >= 7, 1L, 0L),
         conservative = if_else(polviews >= 5, 1L, 0L)) |> 
  drop_na()
```

```{r}
table(conservative = d$conservative, weekly = d$weekly)
```

1.  What is the probability that a person attends religious services weekly?\
    ... and
2.  What is the probability that a person does not attend religious services weekly?

reorganize a bit like we did in class:

```{r}
d <- d |> mutate(weekly = if_else(attend >= 7, 1L, 0L),
          attends = if_else(weekly == 1,"Weekly","Not Weekly"),
          conservative = if_else(polviews >= 5, 1L, 0L),
          polview_chr = if_else(conservative==1,"Conservative","Not Conservative")) |> relocate(attends,weekly,polview_chr,conservative)
```

```{r}
tabyl(d, weekly)
```

1: Pr(W) = 22.237%, and 2: Pr(W\^c) = 77.763

3.  What is the probability a person is conservative, given they attend religious services weekly?

```{r}
tabyl(d, attends, polview_chr) |> 
  adorn_percentages(denominator = "row")
```

Pr(C \| W) = 48.29%

4.  Probability person attends weekly given conservative?

```{r}
tabyl(d, attends, polview_chr) |> 
  adorn_percentages(denominator = "col")
```

Pr(W \| C) = 32.65%

5.  Probability someone is conservative given they don't attend weekly?

Pr(C \| W\^c) = 28.48%

6.  Probability someone attends weekly and is conservative?

```{r}
tabyl(d, attends, polview_chr) |> 
  adorn_percentages(denominator = "all")
```

Pr(W and C) = 10.74%

7.  Probability someone doesn't attend weekly and they are not conservative?

Pr(W\^c and C\^c) = 55.62%

8.  Assuming W and C are independent, what is the joint probability of W and C?

```{r}
tabyl(d, polview_chr)
```

$Pr(W) * Pr(C) = 22.237\%*32.89\% = 7.313\%$

## 7.1.2

```{r}
d2 <- gss18 |> 
  select(astrosci, condrift) |> 
  haven::zap_missing() |> 
  haven::zap_labels() |> 
  drop_na() |> 
  mutate(astro_believe = if_else(astrosci == 1, 1L, 0L),
         astro_believe_chr = if_else(astrosci == 1, "Believes Astrology", "No Astrology"),
         condrift_believe = if_else(condrift == 1, 1L, 0L),
         condrift_believe_chr = if_else(condrift == 1, "Believes Continents Move", "No Continents Move")) |>
  relocate(astro_believe,astro_believe_chr,condrift_believe,condrift_believe_chr)
```

```{r}
table(astro_believe = d2$astro_believe, condrift_believe = d2$condrift_believe)
```

1.  What is the probability that a person believes in astrology?\
    ... and
2.  What is the probability that a person does not believe in astrology?

```{r}
tabyl(d2, astro_believe_chr)
```

1: Pr(A) = 6.73%, and 2: Pr(A\^c) = 93.27%

3.  What is the probability a person believes continental drift, given they believe in astrology?

```{r}
tabyl(d2, astro_believe_chr, condrift_believe_chr) |> 
  adorn_percentages(denominator = "row")
```

Pr(D \| A) = 80%

4.  Probability person believes in astrology given they believe continental drift?

```{r}
tabyl(d2, astro_believe_chr, condrift_believe_chr) |> 
  adorn_percentages(denominator = "col")
```

Pr(A \| D) = 6.2%

5.  Probability someone believes continental drift given they don't believe in astrology?

Pr(D \| A\^c) = 87.32%

6.  Probability someone believes in astrology and continental drift?

```{r}
tabyl(d2, astro_believe_chr, condrift_believe_chr) |> 
  adorn_percentages(denominator = "all")
```

Pr(A and D) = 5.38%

7.  Probability someone doesn't believe in astrology and they don't believe in continental drift?

Pr(A\^c and D\^c) = 11.83%

8.  Assuming A and D are independent, what is the joint probability of A and D?

```{r}
tabyl(d2, condrift_believe_chr)
```

$Pr(A) * Pr(D) = 6.73\%* 86.83\% = 5.84\%$

## 7.2.1

```{r}
# Values from contingency tables considered earlier
pr_w_given_c <- .3265
pr_w_given_not_c <- .1713

# 1. Difference in Probabilities
pr_w_given_c - pr_w_given_not_c

# 2. Relative Risk Ratio
pr_w_given_c / pr_w_given_not_c

# 3. Odds Ratio
odds_w_c <- pr_w_given_c / (1 - pr_w_given_c)
odds_w_not_c <- pr_w_given_not_c / (1 - pr_w_given_not_c)
odds_w_c / odds_w_not_c

# 4. Log Odds Ratio
log(odds_w_c / odds_w_not_c)
```

## 7.2.2

```{r}
boot_risk <- d |> 
  mutate(conservative_cat = if_else(conservative == 1L, "conservative", "other")) |> 
  mutate(weekly_cat = if_else(weekly == 1L, "weekly", "not_weekly")) |> 
  specify(weekly_cat ~ conservative_cat, success="weekly") |> 
  generate(reps = 1e4, type = "bootstrap") |> 
  calculate(stat = "ratio of props", order = c("conservative", "other"))

ci_risk <- get_confidence_interval(boot_risk)

boot_risk |> 
  visualize() +
  shade_ci(ci_risk)+
  ggtitle("Relative Risk Ratio of Pr(W | C), Pr(W | C^c)")
```

```{r}
boot_odds <- d |> 
  mutate(conservative_cat = if_else(conservative == 1L, "conservative", "other")) |> 
  mutate(weekly_cat = if_else(weekly == 1L, "weekly", "not_weekly")) |> 
  specify(weekly_cat ~ conservative_cat, success="weekly") |> 
  generate(reps = 1e4, type = "bootstrap") |> 
  calculate(stat = "odds ratio", order = c("conservative", "other"))

ci_odds <- get_confidence_interval(boot_odds)

boot_odds |> 
  visualize() +
  shade_ci(ci_odds)+
  ggtitle("Odds Ratio of Pr(W | C), Pr(W | C^c)")
```

## 7.2.3

```{r}
# Values from contingency tables considered earlier
pr_a_given_d <- .062
pr_a_given_not_d <- .8978

# 1. Difference in Probabilities
pr_a_given_d - pr_a_given_not_d

# 2. Relative Risk Ratio
pr_a_given_d / pr_a_given_not_d

# 3. Odds Ratio
odds_a_d <- pr_a_given_d / (1 - pr_a_given_d)
odds_a_not_d <- pr_a_given_not_d / (1 - pr_a_given_not_d)
odds_a_d / odds_a_not_d

# 4. Log Odds Ratio
log(odds_a_d / odds_a_not_d)
```

## 7.2.4

```{r}
boot_risk_2 <- d2 |> 
  specify(astro_believe_chr ~ condrift_believe_chr, success="Believes Astrology") |> 
  generate(reps = 1e4, type = "bootstrap") |> 
  calculate(stat = "ratio of props", order = c("Believes Continents Move", "No Continents Move"))

ci_risk_2 <- get_confidence_interval(boot_risk_2)

boot_risk_2 |> 
  visualize() +
  shade_ci(ci_risk_2)+
  ggtitle("Relative Risk Ratio of Pr(A | D), Pr(A | D^c)")
```

```{r}
boot_odds_2 <- d2 |> 
  specify(astro_believe_chr ~ condrift_believe_chr, success="Believes Astrology") |> 
  generate(reps = 1e4, type = "bootstrap") |> 
  calculate(stat = "odds ratio", order = c("Believes Continents Move", "No Continents Move"))

ci_odds_2 <- get_confidence_interval(boot_odds_2)

boot_odds_2 |> 
  visualize() +
  shade_ci(ci_odds_2)+
  ggtitle("Odds Ratio of Pr(A | D), Pr(A | D^c)")
```
