---
title: "HW1"
author: "Ben Rochford"
format: html
theme: vapor
editor: visual
---

Setup:

```{r}
library(tidyverse)
library(palmerpenguins)
data(mtcars)
```

# Week 1 Assignment

### 1.1

```{r}
typeof(mtcars)  # datatype of mtcars
class(mtcars)  # responsible parent class for mtcars
```

```{r}
nrow(mtcars)  # Number of rows in the mtcars table
ncol(mtcars)  # number of columns
length(mtcars)  # length (== number of columns in this case)
dim(mtcars)  # dimensions of table (#rows, #cols) 
rownames(mtcars)  # titles of the rows (leftmost entry)
colnames(mtcars)  # titles of the columns (topmost entry)
```

### 1.2

```{r}
T <- 123  # T not a reserved keyword and can be reassigned
# TRUE <- 123  # TRUE is a reserved keyword and can't be reassigned
```

### 1.3

```{r}
c(1, FALSE)
c("a", 1)
c(TRUE, 1L)
```

my predictions were really good

### 1.4

```{r}
as.integer(FALSE)
as.integer(TRUE)
```

### 1.5

```{r}
x <- sample(c(TRUE, FALSE), size = 75, replace = TRUE)
str(x)
```

```{r}
sum(x)
mean(x)
sum(x) / length(x)
```

### 1.6

```{r}
mtcars["mpg"]  # Gets a subset mpg dataframe of the mtcars dataframe
mtcars[["mpg"]]  # Gets individual elements from mpg comprising a list
identical(mtcars["mpg"], mtcars$mpg)
identical(mtcars[["mpg"]], mtcars$mpg)
```

\[\] gets a subset of the original datatype, while \[\[\]\] always retrieves individualized elements.

### 1.7

```{r}
letters[17]
letters[seq(2,26,by=2)]
letters[8:12]
```

```{r}
letters[[17]]
# letters[[seq(2,26,by=2)]] error
# letters[[8:12]] error
```

### 1.8

```{r}
letters[18] <- NA
```

We ruined letters

#### Permanently

### 1.9

```{r}
mtcars[mtcars$cyl == 4, ]
mtcars[mtcars$mpg > 23, ]
```

### 1.10

The first error happens because no reference "year" is accessible by R lang search. The second one happens because it finds a reference "year" which is not a vector before finding the one that we really meant, and attempts to run a vector table function on it. It has found the tidyverse one.

### 1.11

```{r}
slice(penguins, seq(2, nrow(penguins), by=2))
```

```{r}
slice(penguins, seq(3, nrow(penguins), by=3))
```

### 1.12

```{r}
filter(penguins, species=="Gentoo", island=="Biscoe", body_mass_g < 5500, body_mass_g > 5000)
```
