---
title: "HW9"
author: "Ben Rochford"
format: html
editor: visual
---

```{r}
#| echo: false
#| results: hide
#| include: false
library(conflicted)
conflicts_prefer(dplyr::filter)
library(tidyverse)
library(gssr)
library(ggthemes)
theme_set(theme_few(base_family = "Optima")) 
```

```{r}
gss18 <- gss_get_yr(2018) 

d <- gss18 |> 
  select(attend, polviews, cappun) |> 
  haven::zap_missing() |> 
  haven::zap_labels() |> 
  mutate(
    weekly = if_else(attend >= 7, 1L, 0L),
    conservative = if_else(polviews >= 5, 1L, 0L),
    cappun = if_else(cappun == 1, 1L, 0L)
    ) |> 
  drop_na() |> 
  select(cappun, weekly, conservative) 
```

Contingency table over cappun and weekly:

```{r}
crosstab1 <- table(cp = d$cappun, weekly = d$weekly)
crosstab1

## column percentages
prop.table(crosstab1, margin = 2)
```

### 9.1.1

-   Express the relationship between `cappun` (Y) and `weekly` (X) as a linear probability model.

```{r}
alpha_linear_911 <- 1050/(587+1050)
beta_linear_911 <- 277/(188+277) - alpha_linear_911
cat(sprintf("p(cappun) = %.3f + %.3fx\n", alpha_linear_911, beta_linear_911))
```

-   Express the relationship between `cappun` (Y) and `weekly` (X) as a logistic regression model.

```{r}
alpha_log_911 <- log(1050/587)
beta_log_911 <- log(277/188) - alpha_log_911
cat(sprintf("logit(cappun) = %.3f + %.3fx\n", alpha_log_911, beta_log_911))
```

### 

Contingency table over cappun, weekly, and conservative:

```{r}
crosstab2 <- table(cp = d$cappun, weekly = d$weekly, conservative = d$conservative)

## column percentages
prop.table(crosstab2, margin = c(2, 3))

as_tibble(crosstab2) |> 
  mutate(cp = if_else(cp == "1", "favors", "opposes")) |> 
  pivot_wider(names_from = cp, values_from = n)
```

### 9.1.2

-   Express the relationship between `cappun` (Y), `weekly` (X_1), and `conservative` (X_2) as a linear probability model.

```{r}
alpha_linear_912 <- 682/(485+682)
beta1_linear_912 <- 123/(115+123) - alpha_linear_912
beta2_linear_912 <- 368/(102+368) - alpha_linear_912
beta3_linear_912 <- 154/(73+154) - alpha_linear_912 - beta1_linear_912 - beta2_linear_912
cat(sprintf("p(cappun) = %.3f + %.3f*weekly + %.3f*cons + %.3f*weekly*cons", alpha_linear_912, beta1_linear_912, beta2_linear_912, beta3_linear_912))
```

-   Express the relationship between `cappun` (Y), `weekly` (X_1), and `conservative` (X_2) as a logistic regression model.

```{r}
alpha_log_912 <- log(682/485)
beta1_log_912 <- log(123/115) - alpha_log_912
beta2_log_912 <- log(368/102) - alpha_log_912
beta3_log_912 <- log(154/73) - alpha_log_912 - beta1_log_912 - beta2_log_912
cat(sprintf("logit(cappun) = %.3f + %.3f*weekly + %.3f*cons + %.3f*weekly*cons", alpha_log_912, beta1_log_912, beta2_log_912, beta3_log_912))
```

### 9.2

Contingency table over (posessing a degree of at least) bachelors, (attending church) weekly, and (being) conservative

```{r}
d92 <- gss18 |> 
  select(degree, attend, polviews) |> 
  haven::zap_missing() |> 
  haven::zap_labels() |> 
  mutate(bachelors = if_else(degree >= 3, 1L, 0L),
         weekly = if_else(attend >= 7, 1L, 0L),
         conservative = if_else(polviews >= 5, 1L, 0L)) |> 
  drop_na()
```

```{r}
crosstab92 <- table(BA=d92$bachelors, weekly=d92$weekly, conservative=d92$conservative)

## column percentages
prop.table(crosstab92, margin = c(2, 3))

as_tibble(crosstab92) |> 
  mutate(BA = if_else(BA == "1", ">= bachelors", "< bachelors")) |> 
  pivot_wider(names_from = BA, values_from = n)
```

-   Express the relationship between `bachelors` (Y), `weekly` (X_1), and `conservative` (X_2) as a linear probability model.

```{r}
alpha_linear_92 <- 384/(859+384)
beta1_linear_92 <- 93/(164+93) - alpha_linear_92
beta2_linear_92 <- 132/(363+132) - alpha_linear_92
beta3_linear_92 <- 88/(152+88) - alpha_linear_92 - beta1_linear_92 - beta2_linear_92
cat(sprintf("p(bachelors) = %.3f + %.3f*weekly + %.3f*cons + %.3f*weekly*cons", alpha_linear_92, beta1_linear_92, beta2_linear_92, beta3_linear_92))
```

-   Express the relationship between `bachelors` (Y), `weekly` (X_1), and `conservative` (X_2) as a logistic regression model.

```{r}
alpha_log_92 <- log(384/859)
beta1_log_92 <- log(93/164) - alpha_log_92
beta2_log_92 <- log(132/363) - alpha_log_92
beta3_log_92 <- log(88/152) - alpha_log_92 - beta1_log_92 - beta2_log_92
cat(sprintf("logit(bachelors) = %.3f + %.3f*weekly + %.3f*cons + %.3f*weekly*cons", alpha_log_92, beta1_log_92, beta2_log_92, beta3_log_92))
```

### 9.3

Contingency table over (whether someone is OK with) extramarital sex, (whether someone is) female, (whether someone says they) gossip

```{r}
d93 <- gss18 |> 
  select(mcsds6, sex, xmarsex) |> 
  haven::zap_missing() |> 
  haven::zap_labels() |> 
  mutate(cheating_cool = if_else(xmarsex >= 3 & xmarsex < 8, 1L, 0L),
         female = if_else(sex == 2, 1L, 0L),
         gossip = if_else(mcsds6 == 1, 1L, 0L)) |> 
  drop_na()
```

```{r}
crosstab93 <- table(xmar_sex_ok=d93$cheating_cool, female=d93$female, gossip=d93$gossip)

## column percentages
prop.table(crosstab93, margin = c(2, 3))

as_tibble(crosstab93) |> 
  mutate(xmar_sex_ok = if_else(xmar_sex_ok == "1", "cheating cool", "not cool")) |> 
  pivot_wider(names_from = xmar_sex_ok, values_from = n)
```

-   Express the relationship between xmar_sex_ok (Y), `female` (X_1), and `gossip` (X_2) as a linear probability model.

```{r}
alpha_linear_93 <- 20/(185+20)
beta1_linear_93 <- 13/(149+13) - alpha_linear_93
beta2_linear_93 <- 22/(118+22) - alpha_linear_93
beta3_linear_93 <- 28/(215+28) - alpha_linear_93 - beta1_linear_93 - beta2_linear_93
cat(sprintf("p(cheating cool) = %.3f + %.3f*female + %.3f*gossip + %.3f*female*gossip", alpha_linear_93, beta1_linear_93, beta2_linear_93, beta3_linear_93))
```

-   Express the relationship between xmar_sex_ok (Y), `female` (X_1), and `gossip` (X_2) as a logistic regression model.

```{r}
alpha_log_93 <- log(20/185)
beta1_log_93 <- log(13/149) - alpha_log_93
beta2_log_93 <- log(22/118) - alpha_log_93
beta3_log_93 <- log(28/215) - alpha_log_93 - beta1_log_93 - beta2_log_93
cat(sprintf("logit(cheating cool) = %.3f + %.3f*female + %.3f*gossip + %.3f*female*gossip", alpha_log_93, beta1_log_93, beta2_log_93, beta3_log_93))
```
